---
title: "Exploratory Data Analysis"
output: 
  github_document:
    pandoc_args: --webtex
always_allow_html: yes
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(plotly, quietly = TRUE)
library(cowplot, quietly = TRUE)
library(scales, quietly = TRUE)
```


## Summary
- Each row represent a session by a user.  
- Each user has only 1 session in the dataset.  
- The data is for 1-year period.  
- ~15% sessions resulted in a purchase.  


## Description of the variables, [data source](https://archive.ics.uci.edu/ml/datasets/Online+Shoppers+Purchasing+Intention+Dataset)

| S No. | Variable | Description |
| ----------- | ----------- | ----------- |
| 1 | `Administrative` | Number of Distinct administrative pages |
| 2 | `Informational` | Number of Distinct Informational pages |
| 3 | `ProductRelated` | Number of Distinct Product Related pages |
| 4 | `Administrative_Duration` | Time(in seconds) spent on Administrative pages |
| 5 | `Informational_Duration` | Time(in seconds) spent on Informational pages |
| 6 | `ProductRelated_Duration` | Time(in seconds) spent on Product Related pages |
| 7 | `BounceRates` | Average bounce rate of all web-pages visited by user. For a web-page its the percentage of people who visit the website from that webpage and left without raising any other request |
| 8 | `ExitRates` | Average exit rate of all web-pages visited by user: For a web-page its the percentage of people who exited the website from that webpage |
| 9 | `PageValues` | Average page value of all web-pages visited by user: For a web-page its the average dollar-value of that page which the user visited before completing the transaction |
| 10 | `SpecialDay` | The closeness of site visitng time to a special day (higher chances of a session resulting in a transaction) |
| 11 | `OperatingSystems` | Operating system used by the user |
| 12 | `Month` | Month of Year |
| 13 | `Browser` | Browser used by the user |
| 14 | `Region` | Geographic region |
| 15 | `TrafficType` | Type of Channel user by the user to arrive at the website |
| 16 | `VisitorType` | Type of the visitor |
| 17 | `Weekend` | Weekend indicator |
| 18 | `Revenue` | Revenue transaction indicator |



```{r loading the data}

X_train <- suppressMessages(read_csv("../data/X_train.csv"))
y_train <- suppressMessages(read_csv("../data/y_train.csv"))

mydata <- cbind(X_train, y_train)
setDT(mydata)

num_vars_1 <- c("Administrative", "Informational", "ProductRelated")
num_vars_2 <- c("Administrative_Duration", "Informational_Duration", "ProductRelated_Duration", 
                "BounceRates", "ExitRates", 
                "PageValues", "ProductRelated")
target <- "Revenue"
cat_vars <- setdiff(names(mydata), c(num_vars_1, num_vars_2, target))

x <- sapply(mydata, function(x) sum(is.na(x)))
col_na <- names(x[x>0])

if (is_empty(col_na)){
  print("There is no missing value in this data set.")
} else{
  print(paste("Variables with NA values:", col_na))
}

plot_1 <- ggplot(mydata, aes(Administrative)) + 
  geom_histogram(binwidth = 1, color = 'white') +
  labs(
    title = "Histograms"
  ) + 
  theme_bw()
plot_2 <- ggplot(mydata, aes(Informational)) + 
  geom_histogram(binwidth = 1, color = 'white') +
  labs(
    title = " "
  ) + theme_bw()
plot_3 <- mydata %>% 
  ggplot(aes(x = ProductRelated, y = stat(count) / sum(count), color=Revenue)) + 
  geom_freqpoly(binwidth = 1) + 
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Frequency polygons of ProductRelated"
  ) + theme_bw()
options(repr.plot.width = 30, repr.plot.height = 6)
plot_grid(plot_1, plot_2, nrow=1)
plot_3
```

> There are over 700 classes in the ProductRelated feature, it is possible to treat this feature as categorical or numerical feature. This needs to be decided during the model and feature selection step of this project.


## Summary of Numeric Variables
```{r numerical_variable_exploration}
quantile_dist <- sapply(num_vars_2, FUN=function(x) {
  print(paste("Mean of", x, "is", round(mydata[, mean(get(x))], digits=3), 
              "and standard deviation is", round(mydata[, sd(get(x))], digits=3)))
  mydata[, round(quantile(get(x), probs=seq(0,1,0.05)), digits=3)]
})
print(quantile_dist)

```

## Percentage of rows for categories of categorical variables
```{r categorical_variable_exploration}
# Percentage of rows for categories of categorical variables

sapply(c(cat_vars, target), FUN=function(x) {
  round(mydata[, table(get(x))]/nrow(mydata), digits=3)*100
})

```



## Target vs Other Variables
```{r Target vs Other Variables}

mydata %>% 
  select(num_vars_2, Revenue) %>% 
  gather("num_vars", "values", -Revenue) %>% 
  ggplot(aes(y=values, x=Revenue)) +
  geom_violin(mapping = aes(fill = Revenue),  show.legend = FALSE) +
  facet_wrap(~num_vars, scales = "free", nrow = 3) +
  labs(
    title = "Distributions of the numerical variables"
  ) + 
  theme_bw()

mydata %>% 
  select(cat_vars, Revenue) %>% 
  gather("num_vars", "values", -Revenue) %>% 
  ggplot(aes(x=values, fill=Revenue)) +
  geom_bar(stat = "count", position = "dodge") +
  facet_wrap(~num_vars, scales  ="free", nrow = 4) +
  labs(
    title = "Distributions of the categorical variables"
  ) + 
  theme_bw()

```














